% !TEX root = top.tex
% above command is so that compilation is always from top.tex
\section{Related Work or Background} \label{sec:related}
Our work in this paper draws on concepts from several lines of past research.

\Paragraph{Virtualization Technology} Machine virtualization technology is
a complex and diverse space. Classical Hypervisors/Virtual Machine Monitors
(VMMs) relied on the fundamental primitive of \emph{Trap-and-Emulate}, where
sensitive instructions in the guest would be trapped by the hardware, and
emulated safely within the VMM using shadow structures for privileged
state~\cite{classic-virt}. In x86 however, not all sensitive instructions are
privileged, meaning they can be trapped, and other techniques must also be used
to enable virtualization. \emph{Full virtualization} of unmodified guest
kernels was enabled through binary translation, where all sensitive
instructions could be translated into privileged instructions.
\emph{Paravirtualization} used modifications to the guest operating systems to
ensure sensitive operations were trapped. Today, modern hardware architectures
include special virtualization instructions which remove the need for binary
translation, and additionally remove the need for performance critical shadow
structures using two-dimensional hardware page tables~\cite{virt-techniques}.
Current virtualization technologies offer a mix of all these techniques,
including binary translation for full nested virtualization using Oracle's
HVX~\cite{hvx}, paravirtualization with Xen~\cite{xen}, and classic
trap-and-emulate utilizing modern hardware extensions and the QEMU x86
emulator~\cite{qemu} within the Linux kernel with KVM~\cite{kvm}. Xen is highly
used within the research community because of its relatively simple
software-only techniques, and KVM is valued for its tight integration with the
Linux kernel.

\Paragraph{Serverless Computing} Serverless computing has become an
increasingly important and desirable platform in the cloud ecosystem. Stemming
from the grand vision of computation as a utility, serverless computing offers
users the ability to run application code directly on a black-box
infrastructure. Serverless has the potential to offer an easy to program,
auto-scaling, cost efficient way to utilize cloud infrastructure for users,
without the need to manage machine provisioning and configuration or service
orchestration~\cite{berkeley-serverless}. Although there exist many popular
industry serverless computing platforms
today~\cite{lambda}\cite{gcf}\cite{azure-cf}\cite{openwhisk}, serverless is
still an active area of research, with many improvements to be
made~\cite{peeking}\cite{trilemma}\cite{steps-back}.

\Paragraph{Cold-start Reduction Efforts} There exists a number of different
techniques which have been proposed in order to reduce cold-start
latencies. One broad technique is to reduce the size and complexity of VMs
which are used for serverless functions. Because serverless functions running
in a cloud infrastructure run on a fairly limited and standard set of physical
machines, the number of specialized drivers and kernel modules required to
support this hardware is vastly smaller than that in a standard distribution.
Additionally, because the function will only need to run a single process, a
number of other optimizations can be made to reduce kernel size and boot time.
These kinds of operating system optimizations have been explored previously
with Unikernels~\cite{unikernels}, the Denali OS~\cite{denali}, and are
represented in Amazon's custom $\mu$-Kernel and their serverless KVM VMM
Firecracker~\cite{firecracker}.

Another technique for reducing VM creation latencies involves the process of
\emph{flash-cloning}, in which new VMs are cloned from existing reference VMs.
This is an analogous process to forking processes within linux. One work which
had success with this technique was the Potemkin Virtual
Honeyfarm~\cite{potemkin}. Lightweight VMs were created as Honeypots for
catching security vulnerabilities in cloud applications. This work successfully
implements flash-cloning within the Xen hypervisor~\cite{xen} along with
copy-on-write memory sharing, but focuses on VM density rather than start-up
latency. We implement flash-cloning within a KVM based infrastructure to more
closely match with industry standard serverless infrastructure, Amazon
Firecracker~\cite{firecracker}, and with the goals of minimizing start-up
latency and utilizing native copy-on-write functionality within the Linux
kernel.

\Paragraph{VM Live Migration} Flash-cloning relates very closely to the more
mature virtualization technology of VM live
migration~\cite{post-copy-migration}\cite{snowflock}. Cloud infrastructure has
long required the ability to efficiently migrate VMs across physical hosts. VM
live migration in general contains a superset of the functionality needed for
flash-cloning, but is focused on efficiently supporting the migration of VM
state across a slow network with minimal interruption to the VM. For serverless
computing infrastructure, we are focused on extremely low latency and therefore
do not want the complex techniques used for live-migration.
